---
title: "Galiciamed Paper"
output: html_document
date: "2023-03-20"
---
```{r setup, include=FALSE}
# Libraries
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library("ggpubr")
library(rstatix)
setwd('d:/Github_repositories/Galiciamed/')
```



# Galiciamed data analysis

In this notebook I analyze the data out of the confocal images after analysis with the ImageJ Macro. The analysis is performed as followed:

1. 


# Data Structure


# Folder Processing Function  

The following function process the files from all c.elegans of an experiment generates a dataframe with the data.

```{r, warning=TRUE}

process_experiment <- function(data_folder){
  
  # Empty DF to save the measurements data and lists for the loaded elements
  data_experiment <- data.frame(matrix(ncol = 8, nrow = 0))
  colnames(data_experiment) <- c("Condition","Sample", "Date", "Worm", "Sum_vol","Sum_IntDen", "Media_Mean", "Media_StdDev")
  samples_list <- c()
  conditions_list <- c()
  dates_list <- c()
  
  # Generation of the data for each sample. Each sample's data is inside a folder, then each folder inside the data folder is processed
  # Each folder with one experiment is named ´CONDITION - DATE´ with " - " as separator
  
  sample_folders <- list.dirs(data_folder, recursive = FALSE)
  for (folder_temp in sample_folders){
    # Extraction of the sample from the folder name
    sample <- last(strsplit(folder_temp, '/')[[1]])
    samples_list <- append(samples_list, sample)
    # Extraction of the condition and date from the sample name via regex
    condition <- substring(sample, 1, regexpr("\\s-\\s", sample)-1)
    if (!is.element(condition, conditions_list)){
      conditions_list <- append(conditions_list, condition)} # not to repeat the condition in the list
    date <- substring(sample, regexpr("\\s-\\s", sample)+3)
    if (!is.element(date, dates_list)){
      dates_list <- append(dates_list, date)}
    
  # Reading of every data file inside the folder. Each xls file represents the measurements in one c.elegans sample image
  # .xls  are tsv. Useful columns are selected.
    
    for (file in dir(folder_temp, pattern = "xls")) {
      temp_file <- read.csv(paste(folder_temp, file, sep = "/"), header = TRUE, sep = "\t", dec = ".", fill = TRUE)[,c(1,2,3,6,7,8)]
      if (class(temp_file[[1]]) == "character") {
        # this ´if´ avoids an error if the file is not read properly or was modified in excel, since the firs column should be a numeric index
        temp_file <- as.data.frame(apply(temp_file, 2, function(x) as.numeric(x)))
        temp_file <- na.omit(temp_file)
      }
  # Since the interesting value is the total expression of smn-1, values of each file are aggregated with the sum of the integrated density. This way, each c.elegant becomes an observation into our dataset
      temp_measurements <-
        tryCatch(
          data.frame(
            condition,
            sample,
            date,
            file,
            sum(temp_file$Volume..nm.3.),
            sum(temp_file$IntDen),
            mean(temp_file$Mean),
            mean(temp_file$StdDev)
          ),
          error = function(e) {print(paste("error in --", folder_temp, "--", file))}
        )
      colnames(temp_measurements) <- colnames(data_experiment)
      data_experiment <- rbind(data_experiment, temp_measurements)
    }
  }
  data_experiment <- data_experiment %>% mutate(Condition = as.factor(Condition), Date = as.factor(Date)) %>% mutate(Condition = relevel(Condition, ref = 'control')) # Having the date & condition as factor
 #print("Processed succesfully")
 return(data_experiment)
}

```

## Check that every folder has the proper name structure
```{r}
every_file <- dir('p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/', recursive = TRUE, include.dirs = FALSE)
every_file[!grepl("\\s-\\s", every_file)]
```


# Generation of plots

## Fig.2

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figura-2-smn-principal/'
Fig.2 <- process_experiment(data_folder)
```


```{r}

Fig.2 %>% mutate(Sum_IntDen = Sum_IntDen/1e7) %>% 
  ggplot(aes(x = Condition,y = Sum_IntDen, fill = as.factor(Condition))) + 
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(aes(fill = as.factor(Condition)), size = 1.6, alpha = 0.7, width = 0.15) +
  scale_y_continuous(breaks=seq(0, 1.45, 0.25)) +
  labs(x = "Condition", y = "Integrated Density (a.u)", title = "Fig.2 - Total Integrated density per C.elegans") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) +
  stat_compare_means(
    label = "p.signif",
    method = "t.test",
    ref.group = "control",
    label.y = 1.4
  )

ggsave("output/Fig.2.png", device = 'png')
ggsave("output/Fig.2.svg", device = 'svg')
```
#OLD
# Representación de las condiciones mutantes respecto a la suma de los controles

```{r fig.width=13}
#unique(medidas$Condition)
medidas %>%
  filter() %>%
  ggplot(aes(x = Condition,y = Sum_IntDen, fill = as.factor(Condition))) +
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(aes(fill = as.factor(Condition)), size = 1, alpha = 0.6, width = 0.2) +
  labs(x = "Condition", y = "Integrated Density") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none") +
  stat_compare_means(method = 'anova' ,label.y = max(medidas$Sum_IntDen)*1.15) +
  stat_compare_means(
    label = "p.signif",
    method = "t.test",
    ref.group = "control",
    label.y = max(medidas$Sum_IntDen)*1.05
  )

```
Algunas condiciones resultan significativas respecto a todos los controles, aunque la distribución de controles la incluye -salvo asm-3-. Para evitar esto vamos a normalizar cada gusano con respecto a su control del dia.

# Representación de las condiciones mutantes relativas a la media de expresion de los controles del dia.

### Hallo el factor por el que se va a normalizar cada muestra de un dia y realizo la normalización.
```{r}

normalization_factor <-
  medidas %>% filter(Condition == 'control') %>% group_by(Fecha) %>% summarise(
    control_mean_Sum_vol = mean(Sum_vol),
    control_median_Sum_vol = median(Sum_vol),
    control_sd_Sum_vol = sd(Sum_vol),
    control_mad_Sum_vol = mad(Sum_vol)
  )
```
### Estandarizo los datos a el control del dia
```{r}
medidas <-
  medidas %>% right_join(normalization_factor, by = "Fecha") %>% mutate(
    norm_mean_Sum_vol = (Sum_vol - control_mean_Sum_vol)/control_sd_Sum_vol,
    norm_median_Sum_vol = (Sum_vol - control_median_Sum_vol)/control_mad_Sum_vol
  )
```
### Calculo de los one-sample t.test
Quiero comprobar que cada una de las adquisiciones son de media diferente a cero mediante un one-sample t-test
```{r}

stat.test <- medidas %>%
  filter(Condition != 'control') %>%
  group_by(Condition, Fecha) %>%
  t_test(norm_mean_Sum_vol ~ 1, mu = 0) %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Fecha") %>%
  mutate(y.position = 16) 

```

```{r fig.width=14}
stat.test$x <- c(0.9, 1.2, 1.7, 1.90, 2.1, 2.3, 2.8, 3.2, 3.8, 4, 4.2, 4.8, 5.2, 5.8, 6, 6.2, 6.8, 7, 7.2, 7.7, 7.9, 8.1, 8.3, 8.8, 9, 9.2, 9.8, 10, 10.2, 10.8, 11.2)

medidas %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition,y = norm_mean_Sum_vol, fill = Fecha)) +
  geom_boxplot( outlier.shape = NA, width = 0.6) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.6), size =1, alpha=0.6) +
  labs(x = "Condition", y = "Integrated Density", title = "Normalized change in expression with respect own control") +
  theme_bw(base_size = 15) +
  stat_pvalue_manual(stat.test, label = "p.adj.signif", x = "x", xmax=NULL)

```




