---
title: "Galiciamed Paper"
output: html_document
date: "2023-03-20"
---
```{r setup, include=FALSE}
# Libraries
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("ggpubr")
library("rstatix")
setwd('d:/Github_repositories/Galiciamed/')
```



# Galiciamed Notebook Info

In this notebook I analyze the data out of the confocal images after analysis with the ImageJ Macro. Every image from C.elegans has a .xls (.tsv) file with all objects detected via image analysis. Each of these objects represent *smn-1* expression in the worm. Interested magnitude is *Fluorescent Integrated Density*. In order to obtain a representative value for the worm, all Fluorescent Integrated Density is sum up, interpreting this value as the total expression of *smn-1* inside the region of interest.

## Data Storage Structure

Every experiment is stored in an independent folder. Every condition of that experiment is stored in a subfolder

## Data Processing

The processing of the data is achieved as followed:

1. A folder with all data to be ploted is selected
2. From the path and name of every datafile in that folder, the sample, condition, date and worm number are extracted via Regex
3. The total *Fluorescent Integrated Density* of the worm is sum up and inserted in a dataframe as new row

The former is defined in a function ´process_experiment()´

## Data Standarization

The SMN-1 expression of the candidate genes in Fig.3 and Fig.4 is relative to its control. The control sample is acquired in the same day under same conditions. To plot the relative expression, each sample is standardized to its control sample via substracting the control mean and dividing by the control standard deviation.

## Statistical Significance

XXX

# Folder Processing Function  

The following function process the files from all C.elegans of an experiment generates a dataframe with the data.

```{r, warning=TRUE}

process_experiment <- function(data_folder){
  
  # Empty DF to save the measurements data and lists for the loaded elements
  data_experiment <- data.frame(matrix(ncol = 8, nrow = 0))
  colnames(data_experiment) <- c("Condition","Sample", "Date", "Worm", "Sum_vol","Sum_IntDen", "Media_Mean", "Media_StdDev")
  samples_list <- c()
  conditions_list <- c()
  dates_list <- c()
  
  # Generation of the data for each sample. Each sample's data is inside a folder, then each folder inside the data folder is processed
  # Each folder with one experiment is named ´CONDITION - DATE´ with " - " as separator
  
  sample_folders <- list.dirs(data_folder, recursive = FALSE)
  for (folder_temp in sample_folders){
    # Extraction of the sample from the folder name
    sample <- last(strsplit(folder_temp, '/')[[1]])
    samples_list <- append(samples_list, sample)
    # Extraction of the condition and date from the sample name via regex
    condition <- substring(sample, 1, regexpr("\\s-\\s", sample)-1)
    if (!is.element(condition, conditions_list)){
      conditions_list <- append(conditions_list, condition)} # not to repeat the condition in the list
    date <- substring(sample, regexpr("\\s-\\s", sample)+3)
    if (!is.element(date, dates_list)){
      dates_list <- append(dates_list, date)}
    
  # Reading of every data file inside the folder. Each xls file represents the measurements in one c.elegans sample image
  # .xls  are tsv. Useful columns are selected.
    
    for (file in dir(folder_temp, pattern = "xls")) {
      temp_file <- read.csv(paste(folder_temp, file, sep = "/"), header = TRUE, sep = "\t", dec = ".", fill = TRUE)[,c(1,2,3,6,7,8)]
      if (class(temp_file[[1]]) == "character") {
        # this ´if´ avoids an error if the file is not read properly or was modified in excel, since the firs column should be a numeric index
        temp_file <- as.data.frame(apply(temp_file, 2, function(x) as.numeric(x)))
        temp_file <- na.omit(temp_file)
      }
  # Since the interesting value is the total expression of smn-1, values of each file are aggregated with the sum of the integrated density. This way, each c.elegant becomes an observation into our dataset
      temp_measurements <-
        tryCatch(
          data.frame(
            condition,
            sample,
            date,
            file,
            sum(temp_file$Volume..nm.3.),
            sum(temp_file$IntDen),
            mean(temp_file$Mean),
            mean(temp_file$StdDev)
          ),
          error = function(e) {print(paste("error in --", folder_temp, "--", file))}
        )
      colnames(temp_measurements) <- colnames(data_experiment)
      data_experiment <- rbind(data_experiment, temp_measurements)
    }
  }
  data_experiment <- data_experiment %>% mutate(Condition = as.factor(Condition), Date = as.factor(Date)) %>% mutate(Condition = relevel(Condition, ref = 'control')) # Having the date & condition as factor
 #print("Processed succesfully")
 return(data_experiment)
}

```

## Check that every folder has the proper name structure
```{r}
every_file <- dir('p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/', recursive = TRUE, include.dirs = FALSE)
every_file[!grepl("\\s-\\s", every_file)]
```


# Plots

## Fig.2

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figura-2-smn-principal/'
Fig.2 <- process_experiment(data_folder)
```


```{r message=FALSE}

Fig.2 %>% mutate(Sum_IntDen = Sum_IntDen/1e7) %>% 
  ggplot(aes(x = Condition,y = Sum_IntDen, fill = as.factor(Condition))) + 
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(aes(fill = as.factor(Condition)), size = 1.6, alpha = 0.7, width = 0.15) +
  scale_y_continuous(breaks=seq(0, 1.45, 0.25)) +
  labs(x = "Condition", y = "Integrated Density (a.u)", title = "Fig.2 - Total Integrated density per C.elegans") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) +
  stat_compare_means(
    label = "p.signif",
    method = "t.test",
    ref.group = "control",
    label.y = 1.4
  )

# ggsave("Fig.2.png", device = 'png', path = 'output images/')
# ggsave("Fig.2.svg", device = 'svg', path = 'output images/')
# write_csv(Fig.2, "output data/Fig.2_data.csv")
```




## Fig.3

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-3-11genes-principal/'
Fig.3 <- process_experiment(data_folder)
```

### Data Standarization

```{r}

standarization_factors <-
  Fig.3 %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.3 <-
  Fig.3 %>% right_join(standarization_factors, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

```
### Statistical Test

```{r}

stat.test <- Fig.3 %>%
  filter(Condition != 'control') %>%
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0) %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Date") %>%
  mutate(y.position = 16) 

```


#OLD


### Calculo de los one-sample t.test
Quiero comprobar que cada una de las adquisiciones son de media diferente a cero mediante un one-sample t-test
```{r}

stat.test <- medidas %>%
  filter(Condition != 'control') %>%
  group_by(Condition, Fecha) %>%
  t_test(norm_mean_Sum_vol ~ 1, mu = 0) %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Fecha") %>%
  mutate(y.position = 16) 

```

```{r fig.width=14}
stat.test$x <- c(0.9, 1.2, 1.7, 1.90, 2.1, 2.3, 2.8, 3.2, 3.8, 4, 4.2, 4.8, 5.2, 5.8, 6, 6.2, 6.8, 7, 7.2, 7.7, 7.9, 8.1, 8.3, 8.8, 9, 9.2, 9.8, 10, 10.2, 10.8, 11.2)

medidas %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition,y = norm_mean_Sum_vol, fill = Fecha)) +
  geom_boxplot( outlier.shape = NA, width = 0.6) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.6), size =1, alpha=0.6) +
  labs(x = "Condition", y = "Integrated Density", title = "Normalized change in expression with respect own control") +
  theme_bw(base_size = 15) +
  stat_pvalue_manual(stat.test, label = "p.adj.signif", x = "x", xmax=NULL)

```




