---
title: "Galiciamed Paper"
output: html_document
date: "2023-03-20"
---
```{r setup, include=FALSE}
# Libraries
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("ggpubr")
library("rstatix")
setwd('d:/Github_repositories/Galiciamed/')
# setwd('C:/Users/Usuario/Documents/GitHub/Galiciamed')

```



# Galiciamed Notebook Info

In this notebook I analyze the data out of the confocal images after analysis with the ImageJ Macro. Every image from C.elegans has a .xls (.tsv) file with all objects detected via image analysis. Each of these objects represent *smn-1* expression in the worm. Interested magnitude is *Fluorescent Integrated Density*. In order to obtain a representative value for the worm, all Fluorescent Integrated Density is sum up, interpreting this value as the total expression of *smn-1* inside the region of interest.

## Data Storage Structure

Every experiment is stored in an independent folder. Every condition of that experiment is stored in a subfolder with name structure *CONDITION - DATE*. Note that all control folders have been renamed to *control - date* and some other folders have been renamed due to inconsistency in the naming.

## Data Processing

The processing of the data is achieved as followed:

1. A folder with all data to be ploted is selected
2. From the path and name of every datafile in that folder, the sample, condition, date and worm number are extracted via Regex
3. The total *Fluorescent Integrated Density* of the worm is sum up and inserted in a dataframe as new row

The former is defined in a function ´process_experiment()´

## Data Standarization relative to control

The SMN-1 expression of the candidate genes in Fig.3 and Fig.4 is relative to its control. The control sample is acquired in the same day under same conditions. To plot the relative expression, each sample is standardized to its control sample via substracting the control mean and dividing by the control standard deviation.

## Statistical Significance

For Fig.2 and Fig.5 a single two-sample t-test between populations is applied.
  
To test statistically the candidate genes, since they are standarized to its own control, a one-sample t-test is performed with Null Hypothesis: candidate gene mean expression is equal to zero and Alternative Hypothesis: The mean is *greater* than zero. The p-values are *Bonferroni* corrected, see https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6099145/ 

# Folder Processing Function  

The following function process the files from all C.elegans of an experiment generates a dataframe with the data.

```{r, warning=TRUE}

process_experiment <- function(data_folder){
  
  # Empty DF to save the measurements data and lists for the loaded elements
  data_experiment <- data.frame(matrix(ncol = 8, nrow = 0))
  colnames(data_experiment) <- c("Condition","Sample", "Date", "Worm", "Sum_vol","Sum_IntDen", "Media_Mean", "Media_StdDev")
  samples_list <- c()
  conditions_list <- c()
  dates_list <- c()
  
  # Generation of the data for each sample. Each sample's data is inside a folder, then each folder inside the data folder is processed
  # Each folder with one experiment is named ´CONDITION - DATE´ with " - " as separator
  
  sample_folders <- list.dirs(data_folder, recursive = FALSE)
  for (folder_temp in sample_folders){
    # Extraction of the sample from the folder name
    sample <- last(strsplit(folder_temp, '/')[[1]])
    samples_list <- append(samples_list, sample)
    # Extraction of the condition and date from the sample name via regex
    condition <- substring(sample, 1, regexpr("\\s-\\s", sample)-1)
    if (!is.element(condition, conditions_list)){
      conditions_list <- append(conditions_list, condition)} # not to repeat the condition in the list
    date <- substring(sample, regexpr("\\s-\\s", sample)+3)
    if (!is.element(date, dates_list)){
      dates_list <- append(dates_list, date)}
    
  # Reading of every data file inside the folder. Each xls file represents the measurements in one c.elegans sample image
  # .xls  are tsv. Useful columns are selected.
    
    for (file in dir(folder_temp, pattern = "xls")) {
      temp_file <- read.csv(paste(folder_temp, file, sep = "/"), header = TRUE, sep = "\t", dec = ".", fill = TRUE)[,c(1,2,3,6,7,8)]
      if (class(temp_file[[1]]) == "character") {
        # this ´if´ avoids an error if the file is not read properly or was modified in excel, since the firs column should be a numeric index
        temp_file <- as.data.frame(apply(temp_file, 2, function(x) as.numeric(x)))
        temp_file <- na.omit(temp_file)
      }
  # Since the interesting value is the total expression of smn-1, values of each file are aggregated with the sum of the integrated density. This way, each c.elegant becomes an observation into our dataset
      temp_measurements <-
        tryCatch(
          data.frame(
            condition,
            sample,
            date,
            file,
            sum(temp_file$Volume..nm.3.),
            sum(temp_file$IntDen),
            mean(temp_file$Mean),
            mean(temp_file$StdDev)
          ),
          error = function(e) {print(paste("error in --", folder_temp, "--", file))}
        )
      colnames(temp_measurements) <- colnames(data_experiment)
      data_experiment <- rbind(data_experiment, temp_measurements)
    }
  }
  data_experiment <- data_experiment %>% mutate(Condition = as.factor(Condition), Date = as.factor(Date)) %>% mutate(Condition = relevel(Condition, ref = 'control')) # Having the date & condition as factor
 #print("Processed succesfully")
 return(data_experiment)
}

```

## Check that every folder has the proper name structure
```{r}
every_file <- dir('p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/', recursive = TRUE, include.dirs = FALSE)
every_file[!grepl("\\s-\\s", every_file)]
```


# Plots

## Fig.2

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figura-2-smn-principal/'
Fig.2 <- process_experiment(data_folder)
```


```{r message=FALSE}

Fig.2 %>% mutate(Sum_IntDen = Sum_IntDen/1e7) %>% 
  ggplot(aes(x = Condition,y = Sum_IntDen, fill = as.factor(Condition))) + 
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(aes(fill = as.factor(Condition)), size = 1.6, alpha = 0.7, width = 0.15) +
  scale_y_continuous(breaks=seq(0, 1.45, 0.25)) +
  labs(x = "Condition", y = "Total Fluorescent Integrated Density (a.u)", title = "Fig.2 - Total Fluorescent Integrated density per C.elegans") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) +
  stat_compare_means(
    label = "p.signif",
    method = "t.test",
    ref.group = "control",
    label.y = 1.4
  )

# to save plots and data
ggsave("Fig.2.png", device = 'png', path = 'output images/')
ggsave("Fig.2.svg", device = 'svg', path = 'output images/')
write_csv(Fig.2, "output data/Fig.2_data.csv")
```

## Fig.3

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-3-11genes-principal/'
Fig.3 <- process_experiment(data_folder)
```

### Data Standarization

```{r}

standarization_factors_Fig.3 <-
  Fig.3 %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.3 <-
  Fig.3 %>% right_join(standarization_factors_Fig.3, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

```

### Statistical Test

```{r}

stat.test_Fig.3 <- Fig.3 %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 15.5)
stat.test_Fig.3

```

It is interesting to check the Cohen's d effect size for the t-test, in order to account for the increase in expresion

```{r}
Fig.3 %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')
```

For our gen *asm-3* the magnitude is large, confirming the increase in expresion

### Plot

```{r fig.width=9}

Fig.3 %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = as.factor(Condition))) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
   geom_jitter(aes(fill = as.factor(Condition)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Standarized change in Expression of SMN after RNAi treatment of the 11 candidate genes") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.3, label = "p.adj.signif", x = 'x')
# Note, the position of the significance simbols is in the stat.test DF

# to save plots and data
ggsave("Fig.3.png", device = 'png', path = 'output images/')
ggsave("Fig.3.svg", device = 'svg', path = 'output images/')
write_csv(Fig.3, "output data/Fig.3_data.csv")
```








## Fig.3 Suplementary

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/Figure-3-11genes-supl/'
Fig.3_supl <- process_experiment(data_folder)
```

### Data Standarization

```{r}

standarization_factors_Fig.3_supl <-
  Fig.3_supl %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.3_supl <-
  Fig.3_supl %>% right_join(standarization_factors_Fig.3_supl, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

```

### Statistical Test

```{r}
stat.test_Fig.3_supl <- Fig.3_supl %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 16)
```

It is interesting to check the Cohen's d effect size for the t-test, in order to account for the increase in expresion

```{r}
Fig.3_supl %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')
```

### Plot

```{r fig.width=10}

stat.test_Fig.3_supl$x <- c(1, 1.7,  2,  2.3,  3,  4,  5,  6,  7,  8,  9, 10, 11)

Fig.3_supl %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = as.factor(Date))) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Standarized change in Expression of SMN after RNAi treatment of the 11 candidate genes - Suplementary") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.3_supl, label = "p.adj.signif", x = 'x')
# Note, the position of the significance simbols is in the stat.test DF

# to save plots and data
ggsave("Fig.3_supl.png", device = 'png', path = 'output images/')
ggsave("Fig.3_supl.svg", device = 'svg', path = 'output images/')
write_csv(Fig.3_supl, "output data/Fig.3_supl_data.csv")
```

## Fig.4 A

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easiness
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-4A-clomipramine-gradiente-principal/'
Fig.4A <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Condition, levels = c("control", "clomipramine-30", "clomipramine-60", "clomipramine-100", "clomipramine-120", "clomipramine-150",  "clomipramine-300"))) # ordered factors


### Data Standarization

standarization_factors_Fig.4A <-
  Fig.4A %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.4A <-
  Fig.4A %>% right_join(standarization_factors_Fig.4A, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.4A <- Fig.4A %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0,  alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 15.5)

# Cohen's d

# Fig.4A %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.4A %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Condition)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Treatment with Clomipramine increase the level of SMN.") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.4A, label = "p.adj.signif", x = 'x')
# Note, the position of the significance symbols is in the stat.test DF

# to save plots and data
ggsave("Fig.4A.png", device = 'png', path = 'output images/')
ggsave("Fig.4A.svg", device = 'svg', path = 'output images/')
write_csv(Fig.4A, "output data/Fig.4A_data.csv")
```

## Fig.4 A Suplementary


```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easiness
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/Figure-4A-clomipramine-supl/'
Fig.4A_supl <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Sample)) # to use the sample repetition as the condition
levels(Fig.4A_supl$Condition) <- c("clomipramine-100_rep1", "clomipramine-100_rep2", "clomipramine-100_rep3",
"clomipramine-100_rep4",  "control", "control", "control", "control")  # changing the levels name and grouping the controls


### Data Standarization

standarization_factors_Fig.4A_supl <-
  Fig.4A_supl %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.4A_supl <-
  Fig.4A_supl %>% right_join(standarization_factors_Fig.4A_supl, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.4A_supl <- Fig.4A_supl %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 12)

# Cohen's d

# Fig.4A_supl %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.4A_supl %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Date)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Treatment with Clomipramine 100 uM - Suplementary") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.4A_supl, label = "p.adj.signif", x = 'x')
# Note, the position of the significance simbols is in the stat.test DF

# to save plots and data
ggsave("Fig.4A_supl.png", device = 'png', path = 'output images/')
ggsave("Fig.4A_supl.svg", device = 'svg', path = 'output images/')
write_csv(Fig.4A_supl, "output data/Fig.4A_supl_data.csv")
```

## Fig.4 B

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easiness
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-4B-desipramine-gradiente-principal/'
Fig.4B <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Condition, levels = c("control", "desipramine-15",  "desipramine-50",  "desipramine-250", "desipramine-400", "desipramine-500"))) # ordered factors


### Data Standarization

standarization_factors_Fig.4B <-
  Fig.4B %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.4B <-
  Fig.4B %>% right_join(standarization_factors_Fig.4B, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.4B <- Fig.4B %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0,  alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 15.5)

# Cohen's d

# Fig.4B %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.4B %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Condition)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Treatment with Desipramine increase the level of SMN.") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.4B, label = "p.adj.signif", x = 'x')
# Note, the position of the significance symbols is in the stat.test DF

# to save plots and data
ggsave("Fig.4B.png", device = 'png', path = 'output images/')
ggsave("Fig.4B.svg", device = 'svg', path = 'output images/')
write_csv(Fig.4B, "output data/Fig.4B_data.csv")
```

## Fig.4 B Suplementary

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easinses
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/Figure-4B-desipramine-supl/'
Fig.4B_supl <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Sample)) # to use the sample repetition as the condition
levels(Fig.4B_supl$Condition) <- c("control", "control", "desipramine-400", "desipramine-500_rep1", "desipramine-500_rep2" )  # changing the levels name and grouping the controls


### Data Standarization

standarization_factors_Fig.4B_supl <-
  Fig.4B_supl %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.4B_supl <-
  Fig.4B_supl %>% right_join(standarization_factors_Fig.4B_supl, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.4B_supl <- Fig.4B_supl %>%
  filter(Condition != 'control') %>%  droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 12.5) 

# Cohen's d

# Fig.4B_supl %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.4B_supl %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Date)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Treatment with Clomipramine 100 uM - Suplementary") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.4B_supl, label = "p.adj.signif", x = 'x')
# Note, the position of the significance symbols is in the stat.test DF

# to save plots and data
ggsave("Fig.4B_supl.png", device = 'png', path = 'output images/')
ggsave("Fig.4B_supl.svg", device = 'svg', path = 'output images/')
write_csv(Fig.4B_supl, "output data/Fig.4B_supl_data.csv")
```

## Fig.4 C

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easinses
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-4C-amlodipine-gradiente-principal/'
Fig.4C <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Condition, levels = c("control", "amlodipine-25", "amlodipine-100"))) # ordered factors

### Data Standarization

standarization_factors_Fig.4C <-
  Fig.4C %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.4C <-
  Fig.4C %>% right_join(standarization_factors_Fig.4C, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.4C <- Fig.4C %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0,  alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 8.2)

# Cohen's d

# Fig.4C %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.4C %>%
  filter(Condition != 'control') %>%
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Condition)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Treatment with Amlodipine increase the level of SMN.") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.4C, label = "p.adj.signif", x = 'x')
# Note, the position of the significance symbols is in the stat.test DF

# to save plots and data
ggsave("Fig.4C.png", device = 'png', path = 'output images/')
ggsave("Fig.4.Csvg", device = 'svg', path = 'output images/')
write_csv(Fig.4C, "output data/Fig.4C_data.csv")
```

## Fig.4 C Suplementary

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easiness
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/Figure-4C-amlodipine-supl/'
Fig.4C_supl <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Sample)) # to use the sample repetition as the condition
levels(Fig.4C_supl$Condition) <- c("amlodipine-100_rep1", "amlodipine-100_rep2", "control", "control")  # changing the levels name and grouping the controls

### Data Standarization

standarization_factors_Fig.4C_supl <-
  Fig.4C_supl %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.4C_supl <-
  Fig.4C_supl %>% right_join(standarization_factors_Fig.4C_supl, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.4C_supl <- Fig.4C_supl %>%
  filter(Condition != 'control') %>%  droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 12.5) 

# Cohen's d

# Fig.4C_supl %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.4C_supl %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Date)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Treatment with Amlodipine 100 uM - Suplementary") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.4C_supl, label = "p.adj.signif", x = 'x')
# Note, the position of the significance simbols is in the stat.test DF

# to save plots and data
ggsave("Fig.4C_supl.png", device = 'png', path = 'output images/')
ggsave("Fig.4C_supl.svg", device = 'svg', path = 'output images/')
write_csv(Fig.4C_supl, "output data/Fig.4C_supl_data.csv")
```


## Fig.5

```{r, warning=FALSE}
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-5-clomipramine-adulto-principal/'
Fig.5 <- process_experiment(data_folder)
```


```{r message=FALSE}

Fig.5 %>% mutate(Sum_IntDen = Sum_IntDen/1e7) %>% 
  ggplot(aes(x = Condition,y = Sum_IntDen, fill = as.factor(Condition))) + 
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(aes(fill = as.factor(Condition)), size = 1.6, alpha = 0.7, width = 0.15) +
  scale_y_continuous(breaks=seq(0, 1.45, 0.25)) +
  labs(x = "Condition", y = "Total Fluorescent Integrated Density (a.u)", title = "Fig.5 -  SMN expression after treatment in adult stage-principal") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) +
  stat_compare_means(
    label = "p.signif",
    method = "t.test",
    ref.group = "control",
    label.y = 0.7
  )

# to save plots and data
ggsave("Fig.5.png", device = 'png', path = 'output images/')
ggsave("Fig.5.svg", device = 'svg', path = 'output images/')
write_csv(Fig.5, "output data/Fig.5_data.csv")
```

## Fig.5 Suplementary

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easiness
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/Figure-5-clomipramine-adulto-supl/'
Fig.5_supl <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Sample)) # to use the sample repetition as the condition
levels(Fig.5_supl$Condition) <- c("clomipramine-100_rep1", "clomipramine-100_rep2", "control", "control")  # changing the levels name and grouping the controls

### Data Standarization

standarization_factors_Fig.5_supl <-
  Fig.5_supl %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.5_supl <-
  Fig.5_supl %>% right_join(standarization_factors_Fig.5_supl, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.5_supl <- Fig.5_supl %>%
  filter(Condition != 'control') %>%  droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 6.2) 

# Cohen's d

# Fig.5_supl %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.5_supl %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Date)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Fig.5 - Suplementary: SMN expression after treatment in adult stage") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.5_supl, label = "p.adj.signif", x = 'x')
# Note, the position of the significance simbols is in the stat.test DF

# to save plots and data
ggsave("Fig.5_supl.png", device = 'png', path = 'output images/')
ggsave("Fig.5_supl.svg", device = 'svg', path = 'output images/')
write_csv(Fig.5_supl, "output data/Fig.5_supl_data.csv")
```




## Fig.6 Suplementary

```{r, warning=FALSE, fig.width=10}
# Every command is in a single chunk for easiness
data_folder <- 'p:/CABD/Lab Manolo Muñoz/Ana Maria Brokate/datos paper Galiciamed/figure-6-BN1158-supl/'
Fig.6_supl <- process_experiment(data_folder) %>%
  mutate(Condition = factor(Sample)) # to use the sample repetition as the condition
levels(Fig.6_supl$Condition) <- c("bn-clo100_rep1", "bn-clo100_rep2", "control", "control")  # changing the levels name and grouping the controls

### Data Standarization

standarization_factors_Fig.6_supl <-
  Fig.6_supl %>% filter(Condition == 'control') %>% group_by(Date) %>% summarise(
    control_mean_Sum_IntDen = mean(Sum_IntDen),
    control_sd_Sum_IntDen = sd(Sum_IntDen),
  )

# Adding the Standarization factors to the Fig. dataset using the Date as reference. Afterwards, calculating the Standarized values

Fig.6_supl <-
  Fig.6_supl %>% right_join(standarization_factors_Fig.6_supl, by = "Date") %>%
  mutate(
    stand_Sum_IntDen = (Sum_IntDen - control_mean_Sum_IntDen) / control_sd_Sum_IntDen
    )

### Statistical Test

stat.test_Fig.6_supl <- Fig.6_supl %>%
  filter(Condition != 'control') %>%  droplevels() %>% 
  group_by(Condition, Date) %>%
  t_test(stand_Sum_IntDen ~ 1, mu = 0, alternative = "greater") %>% adjust_pvalue(method = "bonferroni") %>% add_significance("p.adj") %>%
  add_xy_position(x = "Condition") %>%
  mutate(y.position = 4.5) 

# Cohen's d

# Fig.6_supl %>% group_by(Date) %>% cohens_d(Sum_IntDen ~ Condition, ref.group = 'control')

### Plot

Fig.6_supl %>%
  filter(Condition != 'control') %>% droplevels() %>% 
  ggplot(aes(x = Condition, y = stand_Sum_IntDen, fill = Date)) +
  geom_boxplot( outlier.shape = NA, width = 0.5) +
  geom_point(position = position_jitterdodge(jitter.width = 0.8, dodge.width = 0.5), size =1.2, alpha=0.7) +
  # geom_jitter(aes(fill = as.factor(Date)), size = 1.6, alpha = 0.7, width = 0.15) +
  labs(x = "Condition", y = "Relative change in Integrated Density", title = "Fig.6 - Suplementary: SMN expression en cepa BN1158 after treatment in adult stage") +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none', panel.grid.minor = element_blank(), axis.text.x = element_text(size = 11, face = "italic")) + 
  stat_pvalue_manual(stat.test_Fig.6_supl, label = "p.adj.signif", x = 'x')
# Note, the position of the significance simbols is in the stat.test DF

# to save plots and data
ggsave("Fig.6_supl.png", device = 'png', path = 'output images/')
ggsave("Fig.6_supl.svg", device = 'svg', path = 'output images/')
write_csv(Fig.6_supl, "output data/Fig.6_supl_data.csv")
```

Fin
